Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-28 ms.tex
rotation 2

Our spherically symmetric model neglects the effects of angular momentum on the gas evolution.  However, all galaxies possess some net rotation, resulting in centrifugal forces becoming important at some radius $r_{\rm circ} = l_{r_{s}}^{2}/(GM_{\bullet})$.  Here $l_{r_{s}} = \langle r V_{\phi}\rangle |_{r_s}$ is the stellar specific angular momentum near the stagnation radius, from which most of the accreted mass originates, where $V_{\phi}$ is the stellar azimuthal velocity.

@@

\citet{EmsellemCappellari+:2007a} (their Fig.~2) find that $\lambda_R$ is generally $< 0.1$ on radial scales less that 10 per cent of the galaxy half-light radius and that $\lambda_R$ is a decreasing function of $R$, in which case from equation (\ref{eq:rcirc}) we conclude that $r_{\rm circ} \lesssim 0.01r_{s}$.  This location lies well inside the inner sonic point of the flow, suggesting that angular momentum is only becomes important for thermally stable flows in a region that is causally disconnected from the rest of the flow.  To first order we are thus justifying in neglecting its effects on our quantities such as the mass accretion rate.

@@

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccccccc}
  \hline
  {$M_{\bullet}$} & {$v_{w}$} & {$\Gamma^{(a)}$} & {$r_b^{(b)}$} &
  {$\left(\frac{\dot{M}}{\dot{M}_{\rm edd}}\frac{0.04}{\eta}\right)_{r_{s}}^{(c)}$} & {$\left(\frac{t_{\rm cool}}{t_{\rm ff}}\frac{0.04}{\eta}\right)_{r_s}^{(d)}$}  \\
  \hline
  $    10^{ 6 }$ & 300 & 0.8 & 100 & $ 1.3 \times 10^{ -5 }$ & $ 1.1 \times 10^{ 2 }$ \\
  $    10^{ 6 }$ & 600 & 0.8 & 100 & $ 2.2 \times 10^{ -6 }$ & $ 8.1 \times 10^{ 3 }$ \\
  $    10^{ 6 }$ & 1200 & 0.8 & 100 & $ 3.8 \times 10^{ -7 }$ & $ 3.7 \times 10^{ 5 }$ \\
  $    10^{ 7 }$ & 300 & 0.8 & 100 & $ 1.0 \times 10^{ -4 }$ & 6.5 \\
  $    10^{ 7 }$ & 600 & 0.8 & 100 & $ 7.2 \times 10^{ -6 }$ & $ 2.4 \times 10^{ 3 }$ \\
  $    10^{ 7 }$ & 1200 & 0.8 & 100 & $ 1.3 \times 10^{ -6 }$ & $ 1.2 \times 10^{ 5 }$ \\
  $    10^{ 8 }$ & 300 & 0.8 & 100 & $ 6.6 \times 10^{ -4 }$ & $\mathbf{1.4}$ \\
  $    10^{ 8 }$ & 600 & 0.8 & 100 & $ 2.9 \times 10^{ -5 }$ & $ 5.3 \times 10^{ 2 }$ \\
  $    10^{ 8 }$ & 1200 & 0.8 & 100 & $ 4.0 \times 10^{ -6 }$ & $ 3.8 \times 10^{ 4 }$ \\
  $    10^{ 7 }$ & 300 & 0.8 & 25 & $ 7.8 \times 10^{ -5 }$ & 17 \\
  $    10^{ 7 }$ & 300 & 0.8 & 50 & $ 7.9 \times 10^{ -5 }$ & 12 \\
  $    10^{ 7 }$ & 300 & 0.8 & 200 & $ 1.2 \times 10^{ -4 }$ & $\mathbf{4.7}$ \\
  $    10^{ 7 }$ & 300 & 0.8 & 400 & $ 3.6 \times 10^{ -4 }$ & $\mathbf{1.2}$ \\
  $    10^{ 8 }$ & 300 & 0.8 & 25 & $ 1.9 \times 10^{ -4 }$ & 12 \\
  $    10^{ 8 }$ & 300 & 0.8 & 50 & $ 3.3 \times 10^{ -4 }$ &
  $\mathbf{4.6}$ \\
  $    10^{ 8 \dagger}$ & 300 & 0.8 & 50 & $ 9.1 \times 10^{ -4 }$ &
  $\mathbf{5.6}$ \\
  $    10^{ 8 }$ & 300 & 0.8 & 200 & $ 1.4 \times 10^{ -3 }$ &
  $\mathbf{0.43}$ \\
  \hline
  $    10^{ 6 }$ & 1200 & 0.1 & 100 & $ 5.9 \times 10^{ -8 }$ & $ 1.0 \times 10^{ 6 }$ \\
  $    10^{ 7 }$ & 1200 & 0.1 & 100 & $ 5.1 \times 10^{ -7 }$ & $ 9.4 \times 10^{ 4 }$ \\
  $    10^{ 8 }$ & 1200 & 0.1 & 100 & $ 2.4 \times 10^{ -6 }$ & $ 2.1 \times 10^{ 4 }$ \\
  $    10^{ 8 }$ & 1200 & 0.1 & 50 & $ 2.0 \times 10^{ -6 }$ & $ 3.1 \times 10^{ 4 }$ \\
  $    10^{ 8 }$ & 1200 & 0.1 & 200 & $ 4.2 \times 10^{ -6 }$ & $ 8.4 \times 10^{ 3 }$ \\
  $    10^{ 8 }$ & 1200 & 0.1 & 400 & $ 1.2 \times 10^{ -3 }$ & $\mathbf{4.6}$ \\
  $    10^{ 8 }$ & 600 & 0.1 & 100 & $ 8.0 \times 10^{ -4 }$ & $\mathbf{1.6}$ \\
  $    10^{ 6 }$ & 600 & 0.1 & 25 & $ 1.6 \times 10^{ -6 }$ & $ 1.6 \times 10^{ 3 }$ \\
  $    10^{ 6 }$ & 600 & 0.1 & 50 & $ 1.7 \times 10^{ -5 }$ & $\mathbf{33}$ \\
  $    10^{ 7 }$ & 600 & 0.1 & 25 & $ 8.5 \times 10^{ -6 }$ & $ 4.6 \times 10^{ 2 }$ \\
  $    10^{ 7 }$ & 600 & 0.1 & 50 & $ 4.5 \times 10^{ -5 }$ & $\mathbf{21}$ \\
  $    10^{ 8 }$ & 600 & 0.1 & 25 & $ 2.3 \times 10^{ -5 }$ & $ 2.8 \times 10^{ 2 }$ \\
  $    10^{ 8 }$ & 600 & 0.1 & 50 & $ 6.0 \times 10^{ -5 }$ & 56 \\
  \hline

\label{table:models}  
\end{tabular}
\begin{tablenotes}
\item $^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$
Break radius of stellar Nuker profile.  $^{(c)}$Accretion rate in
Eddington units, normalized to a stellar mass input parameter $\eta =
0.04$.  $^{(d)}$Ratio of cooling timescale to free-fall timescale at
stagnation radius.  Thermally unstable solutions are shown in bold
text.
 {\bf  BDM: (3) make sure all combinations of vw and mbh
  are included. AG: Realized that the solution I ran earlier, with
  cooling was actually a solution that would be thermally unstable
  according to our thermal instability. I am worried that this
  solution may not be representative due to the small break radius.}
\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

Figure~\ref{fig:profiles} shows profiles of the density $\rho(r)$,
temperature $T(r)$, and radial velocity $|v(r)|$, for the cusp
($\Gamma=0.8$) solutions within our grid.  As expected, the gas
density increases towards the SMBH $\rho\propto r^{-n}$ with
$n\simeq1$, i.e. shallower than the $-3/2$ power law for
Bondi accretion. This power law behavior does not extend
through all radii, however, as the gas density profile has a break coincident
with the location of the break in the stellar light profile ($\rb=100$
pc). The temperature profile is relatively flat at large radii, but
increases as $\propto 1/r^{k}$ interior to the sphere of influence,
where $k\lesssim 1$, somewhat shallower than expected for virialized gas within the black hole sphere of influence.
The inwardly directed velocity increases towards the hole with a profile that is somewhat steeper than the local free-fall velocity $v \propto v_{\rm ff}\propto r^{-1/2}$.
% {\bf BDM: I made this up - in general we need
%   more explanation of what the plots are actually showing, with some
%   physical explanation.}
% AG -- things are not quite approaching Bondi on the inner grid. I
% don't have a good physical explanation yet...

>>>>

Our spherically symmetric model neglects the effects of angular momentum on the gas evolution.  However, all galaxies possess some net rotation, resulting in centrifugal forces becoming important at some radius $r_{\rm circ} = l^{2}_{r_{s}}/(GM_{\bullet})$.  Here $l_{r_{s}} = \langle r V_{\phi}\rangle |_{r_s}$ is the stellar specific angular momentum near the stagnation radius, from which most of the accreted mass originates, where $V_{\phi}$ is the stellar azimuthal velocity.

@@

\citet{EmsellemCappellari+:2007a} (their Fig.~2) find that $\lambda_R$ is generally $< 0.1$ on radial scales less that 10 per cent of the galaxy half-light radius and that $\lambda_R$ is a decreasing function of $R$.  From equation (\ref{eq:rcirc}) we conclude that $r_{\rm circ} \lesssim 0.01r_{s}$, a location well inside the inner sonic point of the flow (typically located at $\sim 0.1r_{\rm s}$; Fig.~\ref{fig:profiles}).  This suggests that angular momentum is only becomes important for thermally stable flows in a region that is causally disconnected from the rest of the flow.  To first order we are thus justifying in neglecting its effects on our quantities such as the mass accretion rate.

@@

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccc}
  \hline
  {$M_{\bullet}$} & {$v_{w}$} & {$r_b^{(b)}$} &  {$\left(\frac{\dot{M}}{\dot{M}_{\rm edd}}\frac{0.04}{\eta}\right)_{r_{s}}^{(c)}$} & {$\left(\frac{t_{\rm cool}}{t_{\rm ff}}\frac{0.04}{\eta}\right)_{r_s}^{(d)}$}  \\
  \hline
 & & & $\Gamma^{(a)} = 0.8$ & & & & & \\
\hline
  $    10^{ 6 }$ & 300 & 100 & $ 1.3 \times 10^{ -5 }$ & $ 1.1 \times 10^{ 2 }$ \\
  $    10^{ 6 }$ & 600  & 100 & $ 2.2 \times 10^{ -6 }$ & $ 8.1 \times 10^{ 3 }$ \\
  $    10^{ 6 }$ & 1200  & 100 & $ 3.8 \times 10^{ -7 }$ & $ 3.7 \times 10^{ 5 }$ \\
  $    10^{ 7 }$ & 300 & 100 & $ 1.0 \times 10^{ -4 }$ & 6.5 \\
  $    10^{ 7 }$ & 600  & 100 & $ 7.2 \times 10^{ -6 }$ & $ 2.4 \times 10^{ 3 }$ \\
  $    10^{ 7 }$ & 1200  & 100 & $ 1.3 \times 10^{ -6 }$ & $ 1.2 \times 10^{ 5 }$ \\
  $    10^{ 8 }$ & 300 & 100 & $ 6.6 \times 10^{ -4 }$ & $\mathbf{1.4}$ \\
  $    10^{ 8 }$ & 600  & 100 & $ 2.9 \times 10^{ -5 }$ & $ 5.3 \times 10^{ 2 }$ \\
  $    10^{ 8 }$ & 1200  & 100 & $ 4.0 \times 10^{ -6 }$ & $ 3.8 \times 10^{ 4 }$ \\
  $    10^{ 7 }$ & 300  & 25 & $ 7.8 \times 10^{ -5 }$ & 17 \\
  $    10^{ 7 }$ & 300  & 50 & $ 7.9 \times 10^{ -5 }$ & 12 \\
  $    10^{ 7 }$ & 300  & 200 & $ 1.2 \times 10^{ -4 }$ & $\mathbf{4.7}$ \\
  $    10^{ 7 }$ & 300  & 400 & $ 3.6 \times 10^{ -4 }$ & $\mathbf{1.2}$ \\
  $    10^{ 8 }$ & 300  & 25 & $ 1.9 \times 10^{ -4 }$ & 12 \\
  $    10^{ 8 }$ & 300  & 50 & $ 3.3 \times 10^{ -4 }$ &
  $\mathbf{4.6}$ \\
  $    10^{ 8 \dagger}$ & 300  & 50 & $ 9.1 \times 10^{ -4 }$ &
  $\mathbf{5.6}$ \\
  $    10^{ 8 }$ & 300 & 200 & $ 1.4 \times 10^{ -3 }$ &
  $\mathbf{0.43}$ \\
  \hline
 & & & $\Gamma = 0.1$ & & & & & & \\
  \hline
  $    10^{ 6 }$ & 1200  & 100 & $ 5.9 \times 10^{ -8 }$ & $ 1.0 \times 10^{ 6 }$ \\
  $    10^{ 7 }$ & 1200  & 100 & $ 5.1 \times 10^{ -7 }$ & $ 9.4 \times 10^{ 4 }$ \\
  $    10^{ 8 }$ & 1200 & 100 & $ 2.4 \times 10^{ -6 }$ & $ 2.1 \times 10^{ 4 }$ \\
  $    10^{ 8 }$ & 1200  & 50 & $ 2.0 \times 10^{ -6 }$ & $ 3.1 \times 10^{ 4 }$ \\
  $    10^{ 8 }$ & 1200  & 200 & $ 4.2 \times 10^{ -6 }$ & $ 8.4 \times 10^{ 3 }$ \\
  $    10^{ 8 }$ & 1200  & 400 & $ 1.2 \times 10^{ -3 }$ & $\mathbf{4.6}$ \\
  $    10^{ 8 }$ & 600  & 100 & $ 8.0 \times 10^{ -4 }$ & $\mathbf{1.6}$ \\
  $    10^{ 6 }$ & 600  & 25 & $ 1.6 \times 10^{ -6 }$ & $ 1.6 \times 10^{ 3 }$ \\
  $    10^{ 6 }$ & 600  & 50 & $ 1.7 \times 10^{ -5 }$ & $\mathbf{33}$ \\
  $    10^{ 7 }$ & 600  & 25 & $ 8.5 \times 10^{ -6 }$ & $ 4.6 \times 10^{ 2 }$ \\
  $    10^{ 7 }$ & 600  & 50 & $ 4.5 \times 10^{ -5 }$ & $\mathbf{21}$ \\
  $    10^{ 8 }$ & 600  & 25 & $ 2.3 \times 10^{ -5 }$ & $ 2.8 \times 10^{ 2 }$ \\
  $    10^{ 8 }$ & 600  & 50 & $ 6.0 \times 10^{ -5 }$ & 56 \\
  $    10^{ 8 }$ & 300  & 50 & $ 6.0 \times 10^{ -5 }$ & 56 \\
  \hline
\label{table:models}  
\end{tabular}
\begin{tablenotes}
\item $^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$
Break radius of stellar Nuker profile.  $^{(c)}$Accretion rate in
Eddington units, normalized to a stellar mass input parameter $\eta =
0.04$.  $^{(d)}$Ratio of cooling timescale to free-fall timescale at
stagnation radius.  Thermally unstable solutions are shown in bold
text.
 {\bf  BDM: (3) make sure all combinations of vw and mbh
  are included. AG: Realized that the solution I ran earlier, with
  cooling was actually a solution that would be thermally unstable
  according to our thermal instability. I am worried that this
  solution may not be representative due to the small break radius.}
\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

Figure~\ref{fig:profiles} shows profiles of the density $\rho(r)$,
temperature $T(r)$, and radial velocity $|v(r)|$ (expressed in units of the sound speed), for the cusp
($\Gamma=0.8$) solutions within our grid.  As expected, the gas
density increases towards the SMBH $\rho\propto r^{-n}$ with
$n\simeq1$, i.e. shallower than the $-3/2$ power law for
Bondi accretion. This power law behavior does not extend
through all radii, however, as the gas density profile has a break coincident
with the location of the break in the stellar light profile ($\rb=100$
pc). The temperature profile is relatively flat at large radii, but
increases as $\propto 1/r^{k}$ interior to the sphere of influence,
where $k\lesssim 1$, somewhat shallower than expected for virialized gas within the black hole sphere of influence.
The inwardly directed velocity increases towards the hole with a profile that is somewhat steeper than the local free-fall velocity $v \propto v_{\rm ff}\propto r^{-1/2}$.  The flow near the stagnation radius is subsonic, but becomes supersonic at two critical points, one located at small radii, $r \lesssim 0.1r_{\rm s}$, and the other located at large radii, $r \gtrsim 10^{3}r_{\rm s}$.  In many cases the supersonic transition is instigated by the transition to a steeper stellar density prorfile exterior to the outer Nuker break radius, making our solutions near $r_{\rm s}$ also potentially sensitive to the break radius value.
% {\bf BDM: I made this up - in general we need
%   more explanation of what the plots are actually showing, with some
%   physical explanation.}
% AG -- things are not quite approaching Bondi on the inner grid. I
% don't have a good physical explanation yet...
